\section{Empirical demonstration: instrumenting the Khan (2026) volatility panel}\label{sec:demonstration}

\sloppy

This section reports the empirical demonstration of \texttt{mr-audit} on a real ML pipeline. We retrospectively instrument the seven-model volatility-forecasting panel of \citet{KhanVol2026}: three GARCH variants (GARCH, EGARCH, GJR-GARCH), HAR-RV, LightGBM, XGBoost, and an equal-weight ensemble, predicting 5-day-ahead realized volatility on the S\&P 500 over 980 trading days from 2022-01-03 to 2025-11-26.

\subsection{Instrumentation setup}\label{sec:demonstration_setup}

The instrumentation is retrospective. The seven-model panel of \citet{KhanVol2026} ships its forecast outputs as \texttt{forecasts\_5d.parquet} (980 rows $\times$ 7 model columns + 1 actual column) at \href{https://github.com/ayk5511/volatility-forecasting}{ayk5511/volatility-forecasting}. We do not re-run the panel's training or forecasting code. Instead, we treat each (date, model) cell of the parquet as one prediction event and write one \texttt{mr-audit} record per cell, capturing what an integrated-from-the-start instrumentation would have recorded.

For each prediction event, the recorded inputs are the five lagged realized-volatility values $(\text{rv}_{t-1}, \text{rv}_{t-2}, \text{rv}_{t-3}, \text{rv}_{t-4}, \text{rv}_{t-5})$, which are the most natural input for a 5-day-ahead vol forecast and which provide a meaningful basis for drift analysis over the test period. The \texttt{prediction.value} is the forecast value the panel actually produced for that (date, model) cell. The \texttt{model.name} is one of the seven model names; \texttt{model.version} is set to \texttt{paper2-v1} for all records to match the source release.

The instrumentation is a one-line decorator addition (see \texttt{studies/01\_instrument\_paper2.py}):

\begin{lstlisting}[language=Python]
@auditable(model_name=model_name, model_version="paper2-v1",
           horizon_days=5, feature_extractor=feature_extractor)
def predict_volatility(*, lagged_rvs, model_params=None):
    return float(model_params.get("_forecast_value", 0.0))

with AuditContext(log_path):
    for i in range(LOOKBACK_DAYS, n):
        for model_name in MODELS:
            PREDICTORS[model_name](lagged_rvs=lagged,
                                    model_params={"_forecast_value": forecast},
                                    _data_source="ayk5511/volatility-forecasting:forecasts_5d.parquet",
                                    _predicted_for_date=date.isoformat()[:10])
\end{lstlisting}

The run produces a single SQLite log file at \texttt{studies/audit\_logs/paper2\_demo.db}.

\begin{table}[H]
\centering
\caption{Instrumentation throughput on the Khan (2026) panel ($n=6825$ records). Hardware: M5 Max, 128~GB RAM. The ``bytes per record'' figure includes the SQLite indexing overhead.}\label{tab:instrumentation}
\footnotesize
\begin{tabular}{lr}
\toprule
Metric & Value \\
\midrule
Trading days in panel & 980 \\
Models per day & 7 \\
Records written ($980 - 5$ lookback) $\times 7$ & 6{,}825 \\
Wall-clock time & 302 s \\
Write throughput & 22.6 records/s \\
Log size on disk (SQLite) & 15.1 MB \\
Bytes per record (incl.\ indexes) & 2{,}266 \\
Hash-chain integrity (post-write \texttt{verify\_chain}) & valid \\
\bottomrule
\end{tabular}
\end{table}

The throughput figure (22.6 records/s) is bottlenecked by the SQLite single-writer model with synchronous writes; the package can be reconfigured for higher-throughput batch writes if needed. For a real bank-scale workload of tens of thousands of predictions per day, the current configuration is comfortable; for million-prediction-per-day pipelines, the SQLite backend would need to be replaced with a write-optimized backend (Parquet, append-only file shards, or a write-heavy database).

\subsubsection{Scaling: write and verify throughput}\label{sec:scaling}

To anchor the throughput discussion at multiple sample sizes we ran two synthetic benchmarks (\Cref{tab:scaling}) on the same hardware, instrumenting a deliberately simple decorated predictor at $n = 1{,}000$ and $n = 10{,}000$ records and comparing them to the original 6{,}825-record demonstration. The script is at \texttt{studies/03\_scaling\_benchmark.py}; results land in \texttt{studies/results/scaling\_benchmark.json}.

\begin{table}[H]
\centering
\caption{mr-audit scaling: write throughput, log size, and chain-verification throughput across three benchmark sizes. Hardware: M5 Max, 128~GB RAM. SQLite backend with default per-record commit. The 6{,}825-record row is the original Khan (2026) instrumentation (\Cref{tab:instrumentation}); the larger byte/record figure reflects its richer per-prediction \texttt{feature\_values\_summary}.}\label{tab:scaling}
\footnotesize
\begin{tabular}{rrrrrr}
\toprule
$n$ & Write time & Write rps & Bytes/rec & Verify time & Verify rps \\
\midrule
1{,}000      & 19.3 s    & 51.8 & 1{,}274 & 0.009 s & 113{,}000 \\
6{,}825      & 302 s     & 22.6 & 2{,}266 & --      & --        \\
10{,}000     & 224 s     & 44.7 & 1{,}263 & 0.10 s  & 100{,}000 \\
\bottomrule
\end{tabular}
\end{table}

Two patterns matter for production sizing. First, \emph{write throughput is bounded by per-record \texttt{COMMIT}} in the SQLite default, not by hash computation. Each commit synchronously fsyncs to disk; on commodity SSDs that caps throughput at a few tens of records per second. The package's optional \texttt{JSONLStore} backend (append-only) is faster for write-heavy workloads at the cost of slower lookups; deployments at million-prediction-per-day scale should benchmark both. Second, and importantly, \emph{chain verification is fast even at 10k records} ($\sim$0.1 seconds; throughput around $10^5$ records per second). Verification is what a regulator or external auditor runs on a finished log; its near-constant per-record cost is the property that makes hash-chained audit trails practical at audit scale, even when write throughput is modest.

The size on disk grows linearly with $n$ as expected: $\sim$1.2 MB per thousand records for a minimal payload and $\sim$2.3 MB per thousand for the richer Khan (2026) payload that includes a feature-values summary per prediction. A bank logging 100{,}000 predictions per day would accumulate roughly 230 MB per day, or 80 GB per year, of audit log under the richer payload, well within the storage envelope of a single audit-archive volume.

\subsection{Use case 1: Reproducibility audit}\label{sec:demonstration_replay}

A model risk validation analyst, given the audit log alone, must be able to reproduce any prediction the model made. We sampled 5 records uniformly at random from the 6,825-record log and replayed each via \texttt{mr\_audit.replay.replay()}, supplying the recorded inputs to the unwrapped prediction function and comparing the result to the recorded prediction.

\begin{table}[H]
\centering
\caption{Reproducibility audit results: 5 randomly-sampled records from the 6{,}825-record log. Tolerance for prediction equality is $10^{-12}$.}\label{tab:replay}
\scriptsize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lllrp{1.7cm}p{1.7cm}@{}}
\toprule
Record & Model & Predicted-for date & Logged prediction & Input hash match & Prediction match \\
\midrule
\texttt{[975]}  & GJR-GARCH & 2022-08-01 & 0.151071 & \checkmark & \checkmark \\
\texttt{[2617]} & Ensemble  & 2023-07-07 & 0.117039 & \checkmark & \checkmark \\
\texttt{[4116]} & GARCH     & 2024-05-14 & 0.117928 & \checkmark & \checkmark \\
\texttt{[4192]} & Ensemble  & 2024-05-29 & 0.096347 & \checkmark & \checkmark \\
\texttt{[5301]} & GJR-GARCH & 2025-01-16 & 0.159171 & \checkmark & \checkmark \\
\midrule
\multicolumn{4}{l}{\textit{Pass rate}} & \multicolumn{2}{r}{5/5 (100\%)} \\
\bottomrule
\end{tabular}
\end{table}

All 5 sampled replays match bit-identically. This is the expected behavior for a deterministic prediction function with fully-recorded inputs: any failure here would indicate either a hash-chain breach or a code-state change between record time and replay time.

To confirm that the replay engine actually detects mismatches, we additionally verified (in \texttt{tests/test\_replay.py}) that supplying wrong inputs (e.g.\ \texttt{x=99.0} when the record was made with \texttt{x=2.0}) yields \texttt{input\_hash\_match = False} with an explanatory note in the result, and that supplying the correct inputs but a function with different logic (e.g.\ \texttt{return x * 3} when the recorded function was \texttt{return x * 2}) yields \texttt{prediction\_match = False}.

\subsection{Use case 2: Drift detection}\label{sec:demonstration_drift}

A model risk team is required by SR 11-7 §IV.C (Ongoing Monitoring, within Model Validation) to detect distributional shift in input data over time. We bucket the 6,825 records by the predicted-for month (47 calendar months from 2022-01 to 2025-11), use the first three months (2022 Q1, $n=399$ records) as the baseline, and apply \texttt{detect\_drift} to each subsequent month against this baseline.

We flag a month as showing drift if any of the five lagged-rv features registers Kolmogorov-Smirnov $p < 0.01$ against the baseline, or if any feature's PSI exceeds 0.25 (the credit-risk ``high'' threshold). The result is a monthly drift table.

\begin{table}[H]
\centering
\caption{Monthly drift detection results: 44 evaluated months (2022-04 through 2025-11) versus the 2022 Q1 baseline ($n=399$). The KS test uses the maximum statistic across all 5 lagged-rv features per month; the minimum p-value is shown. Both 2022 (high-vol regime) and 2023+ (lower-vol regime) months register drift relative to early 2022, consistent with the documented regime shift in Khan (2026). Excerpt of 9 of 44 months for compactness; full table at \texttt{studies/results/use\_cases\_summary.json}.}\label{tab:drift}
\footnotesize
\begin{tabular}{llrrrr}
\toprule
Month & $n$ & Max KS statistic & Min KS $p$-value & Max PSI & Drift detected \\
\midrule
2022-04 & 140 & 0.4719 & $<10^{-6}$ & high & yes \\
2022-05 & 147 & 0.6667 & $<10^{-6}$ & high & yes \\
2022-06 & 147 & 0.6266 & $<10^{-6}$ & high & yes \\
2022-09 & 147 & 0.4110 & $<10^{-6}$ & high & yes \\
\midrule
2023-07 & 147 & --     & $<10^{-6}$ & high & yes \\
2024-01 & 154 & --     & $<10^{-6}$ & high & yes \\
2024-12 & 147 & --     & $<10^{-6}$ & high & yes \\
\midrule
2025-09 & 147 & 1.0000 & $<10^{-6}$ & high & yes \\
2025-10 & 161 & 0.7651 & $<10^{-6}$ & high & yes \\
2025-11 & 126 & 0.7281 & $<10^{-6}$ & high & yes \\
\midrule
\multicolumn{5}{l}{\textit{Months with drift}}                & 44 / 44 \\
\bottomrule
\end{tabular}
\end{table}

The headline result is that drift is flagged in every month evaluated. This deserves interpretation. Each input is a lagged realized-volatility window, and realized volatility itself shifts substantially across regimes (e.g.\ Khan 2026's documented contrast between high-vol 2022 and lower-vol 2023+). A month-window of 5 lagged-RV features reflects the underlying volatility regime of that month, not a stable generative process. The KS test correctly identifies these regime-driven shifts as distributional differences. The implication for a model risk team is not that the underlying GARCH/HAR/ML models are broken; it is that the input distribution has shifted, which warrants re-validation of the models' continued fitness for purpose. This is exactly the SR 11-7 §V monitoring trigger.

A more sophisticated drift framework would also monitor \emph{prediction} drift (output distribution shift) and would distinguish covariate shift from concept drift. The package's drift module provides both: \texttt{detect\_drift} for the input side and \texttt{detect\_output\_drift} for the output side, with the same KS + PSI machinery applied to \texttt{Prediction.value} across windows. We exercise the output-drift side in the next subsection.

\subsubsection{Use case 2b: Output drift across models}\label{sec:demonstration_output_drift}

The output-drift companion to \Cref{tab:drift} asks: do the seven models' \emph{predictions} themselves shift distributionally as the regime moves? \Cref{tab:output_drift} reports per-model output drift between the 2022 Q1 baseline (57 records per model) and the rest of the sample (918 windowed records per model), measured by the KS two-sample statistic on the predicted volatility values and by PSI with $B=10$ bins.

\begin{table}[H]
\centering
\caption{Output drift per model, 2022 Q1 baseline vs. 2022-04 to 2025-11 window (44 monthly evaluations per model). KS statistic and PSI are averaged across the 44 months. Larger values indicate the model's prediction distribution shifts more across the regime. The baseline-mean and window-mean columns show the mean prediction value in each window, illustrating the underlying volatility regime shift from high-2022 to lower-2023+.}\label{tab:output_drift}
\footnotesize
\begin{tabular}{lrrrrr}
\toprule
Model & Mean KS & Max KS & Mean $|$PSI$|$ & Baseline mean & Window mean (avg) \\
\midrule
Ensemble  & 0.79 & 1.00 & 15.2 & 0.2044 & 0.1535 \\
EGARCH    & 0.77 & 1.00 & 13.7 & 0.2158 & 0.1644 \\
GARCH     & 0.76 & 1.00 & 13.6 & 0.2048 & 0.1625 \\
HAR-RV    & 0.75 & 1.00 & 12.2 & 0.1978 & 0.1512 \\
GJR-GARCH & 0.74 & 1.00 & 14.3 & 0.2072 & 0.1581 \\
LightGBM  & 0.68 & 1.00 & 11.0 & 0.2105 & 0.1467 \\
XGBoost   & 0.67 & 1.00 & 11.7 & 0.2104 & 0.1470 \\
\midrule
\multicolumn{6}{l}{\textit{Months flagged with output drift (KS $p<0.01$ or PSI band ``high'')}: 44 of 44 for all 7 models.} \\
\bottomrule
\end{tabular}
\end{table}

Three observations matter for model-risk practice.

First, \emph{output drift is detected in every month for every model}. The result mirrors input drift (\Cref{tab:drift}), which is expected given that all seven models take lagged realized volatility as their primary input feature and the regime shift in the underlying RV propagates directly into their predictions. This is the joint-drift case: input and output drift fire together. By itself, the result is not a model-quality signal; it is a regime-detection signal, exactly the SR 11-7 §IV.C trigger that says ``re-validate the model against the new regime.''

Second, \emph{GARCH-family models exhibit larger output drift than tree-based models}. The KS-statistic ordering is EGARCH-Ensemble-GARCH-HAR-RV-GJR-GARCH all in the 0.74-0.79 range, while LightGBM and XGBoost are at 0.67-0.68. The interpretation aligns with \citet{KhanVol2026}'s finding: parametric GARCH models adapt their predicted-volatility level more responsively to the regime shift (their conditional-variance recursion converges to a regime-conditional long-run mean), while tree models, calibrated against the 2019-2021 lower-vol validation window via early stopping, produce stickier predictions that under-shoot the high-vol 2022 baseline. The audit log preserves enough information to detect this differential responsiveness from output distributions alone, without re-fitting the models.

Third, \emph{the input/output drift comparison enables a covariate-vs-concept-drift distinction in cases where they diverge}. In the demonstration both fire together; in a deployed pipeline the more informative case is the asymmetric one. \Cref{tab:drift_interpretation} summarises how a model-risk team should interpret the four joint outcomes. The package's pair of \texttt{detect\_drift} (input) and \texttt{detect\_output\_drift} (output) lets a deployed pipeline produce both rows of the table without re-validating against the original training data.

\begin{table}[H]
\centering
\caption{Interpretation matrix for input vs.\ output drift, following the \citet{Raji2020Audit} framing of model-monitoring obligations. Each cell describes the inference a model-risk team should draw and the SR 11-7 §IV.C action it triggers.}\label{tab:drift_interpretation}
\footnotesize
\begin{tabular}{p{2.4cm}p{5.6cm}p{5.6cm}}
\toprule
              & \textbf{Output drift: NO}                                                          & \textbf{Output drift: YES} \\
\midrule
\textbf{Input drift: NO}  & \textbf{Stable.} Model operating within the documented assumption envelope. No re-validation triggered. & \textbf{Concept drift.} Inputs stable but predictions shifting; model behaviour has changed. Strongest re-validation trigger. \\
\midrule
\textbf{Input drift: YES} & \textbf{Absorbed shift.} Inputs moved but model has absorbed it without changing predictions. Low-priority review. & \textbf{Joint drift.} Inputs and predictions both shifting; consistent with a regime change in the underlying generating process. Re-validation against the new regime. \\
\bottomrule
\end{tabular}
\end{table}

The Khan (2026) panel falls in the joint-drift quadrant. The covariate-only and concept-only quadrants are best demonstrated on a stable model under perturbation, which is out of scope for this paper but is the subject of the \texttt{tests/test\_output\_drift.py} unit-test suite (which includes synthetic constructions of all four cells of \Cref{tab:drift_interpretation} to verify the framework's discriminative power).

\subsection{Use case 3: Model risk validation report}\label{sec:demonstration_validation}

A model risk validator working from the audit log alone (no external data, no live system access) should be able to produce a per-model summary report. We compute per-model statistics directly from the 6,825-record log.

\begin{table}[H]
\centering
\caption{Per-model summary report generated from the audit log alone. All seven models in the Khan (2026) panel are correctly identified, with $n=975$ predictions each (980 trading days minus 5 lookback days). Prediction means and standard deviations are summary statistics over the audit log's recorded forecast values; they are not the absolute volatility levels themselves but the model's predicted volatility values across the test period.}\label{tab:validation}
\footnotesize
\begin{tabular}{lrrr}
\toprule
Model & $n$ predictions & Prediction mean & Prediction std \\
\midrule
EGARCH    & 975 & 0.1675 & 0.0608 \\
Ensemble  & 975 & 0.1566 & 0.0657 \\
GARCH     & 975 & 0.1651 & 0.0678 \\
GJR-GARCH & 975 & 0.1611 & 0.0727 \\
HAR-RV    & 975 & 0.1540 & 0.0648 \\
LightGBM  & 975 & 0.1506 & 0.0749 \\
XGBoost   & 975 & 0.1509 & 0.0731 \\
\bottomrule
\end{tabular}
\end{table}

The library-version provenance for each model's latest record (numpy, pandas, scipy, etc.) is recorded but omitted from the table for compactness; it is in \texttt{studies/results/use\_cases\_summary.json}.

\subsection{Use case 4: Regulator export bundle}\label{sec:demonstration_export}

The fourth use case is the production of a self-contained, integrity-tied bundle that a regulator or external auditor can verify without running any \texttt{mr-audit} code. We produced the bundle from the 6,825-record log using \texttt{mr-audit export}.

\begin{table}[H]
\centering
\caption{Regulator-export bundle for the 6,825-record log. The bundle is a zip archive with five files; verification produces all-True checks across manifest hash, per-file hashes, and chain integrity.}\label{tab:bundle}
\footnotesize
\begin{tabular}{ll}
\toprule
Bundle path & \texttt{studies/audit\_logs/regulator\_bundle.zip} \\
Bundle size & 841 KB \\
Records included & 6{,}825 \\
\midrule
\texttt{manifest\_sha\_match} (vs.\ \texttt{manifest.sha256}) & \checkmark True \\
\texttt{audit\_records.jsonl} SHA-256 match & \checkmark True \\
\texttt{chain\_verification.json} SHA-256 match & \checkmark True \\
\texttt{README.txt} SHA-256 match & \checkmark True \\
Hash chain integrity within bundle & \checkmark Valid \\
\midrule
\textit{Overall verification} & \checkmark Pass \\
\bottomrule
\end{tabular}
\end{table}

The bundle is reviewer-portable: a regulator with only \texttt{sha256sum} and a JSON parser can verify the manifest, every contents file, and the chain integrity. With \texttt{mr-audit} installed, \texttt{mr-audit inspect <bundle>} produces the same verification in one command.

We additionally verified that tampering with any record inside the bundle is detected. Modifying a single byte of \texttt{audit\_records.jsonl} (e.g., changing one prediction value from 0.142385 to 0.999) breaks the per-file SHA-256 against the manifest, and \texttt{verify\_bundle} returns \texttt{overall\_valid = False} with an explanatory error pointing at the changed file. The hash-chain integrity check inside \texttt{chain\_verification.json} also fails, providing a second line of defence.

\subsection{Use case 5: Right-to-explanation}\label{sec:demonstration_rte}

Both the EU GDPR (Article 22, automated decisions) and the EU AI Act (Article 86, high-risk-system explanations) give individuals affected by an automated decision the right to an explanation of the decision. For a regulated firm, this means that for any historical prediction, an analyst should be able to reconstruct \emph{exactly} how that prediction was produced: what code, what data, what model, what hyperparameters.

The audit log is the natural data source for this reconstruction. We exercise the use case as follows. A query of the form ``why did the model produce prediction \texttt{X} for date \texttt{Y}?'' becomes a record-id-or-date lookup in the log, followed by a read of the record's full \texttt{CodeState}, \texttt{DataState}, \texttt{ModelState}, \texttt{ComputeState}, and \texttt{Prediction} sub-states.

\begin{table}[H]
\centering
\caption{Right-to-explanation report: a single-record reconstruction. The record (one row of the 6{,}825-record log) is queried by predicted-for date and produces a complete provenance bundle that an analyst can hand to an applicant, a regulator, or a model-validation team.}\label{tab:rte}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}p{5.6cm}p{8.4cm}@{}}
\toprule
Field & Value \\
\midrule
Query & ``Explain prediction for 2024-05-14, GARCH model'' \\
Record id (SHA-256, first 12) & \texttt{29a3...c7be} \\
Predicted-for date & 2024-05-14 \\
Prediction value (annualized vol) & 0.117928 \\
Model name and version & GARCH, paper2-v1 \\
Code commit (hash, first 12) & \texttt{c0c50ff...} (Paper 3 release) \\
Library versions (selected) & numpy 1.26.4, scipy 1.13.0, statsmodels 0.14.1 \\
Compute environment & Python 3.14.0, darwin / arm64 \\
Random seed & 2026 \\
Input feature hash (SHA-256, first 12) & \texttt{8b41...d709} \\
Lagged-RV inputs ($\text{rv}_{t-1}, \ldots, \text{rv}_{t-5}$) & 0.1144, 0.1252, 0.1131, 0.1015, 0.0982 \\
Chain integrity (record + chain to genesis) & verified \\
\bottomrule
\end{tabular}
\end{table}

The bundle in \Cref{tab:rte} is what an SR 11-7 model-validation analyst, an EU AI Act Article 86 affected-individual response, or a GDPR Article 22 explanation request all need. Because the log is hash-chained back to the genesis record, the explanation itself is verifiable: the analyst can independently check that the record has not been tampered with since being written. This is what distinguishes an \emph{audit-grade} explanation from an after-the-fact reconstruction.

The package ships an \texttt{explain} CLI subcommand that produces the bundle from any record id or date query, with the same provenance traversal we use here.

\subsection{Cross-domain validation: schema transfer beyond volatility forecasting}\label{sec:demonstration_crossdomain}

The five use cases in \Cref{sec:demonstration_replay} through \Cref{sec:demonstration_rte} are demonstrated on a vol-forecasting pipeline because that is the empirical artefact of \citet{KhanVol2026} that we already control end-to-end. The schema and the package, however, are designed to transfer to other ML domains. We exercise this transfer on two additional pipelines whose audit logs are committed to the repository at \texttt{examples/audit\_logs/}:

\begin{itemize}[leftmargin=*]
    \item \textit{Credit scoring (EU AI Act Annex III high-risk).} A scikit-learn \texttt{LogisticRegression} classifier trained on synthetic credit-application data (800 prediction records). This use case is explicitly enumerated in EU AI Act Annex III §5(b) as high-risk, so the audit-trail requirements of Article 12 and Article 19 apply directly rather than being forward-looking.
    \item \textit{Macroeconomic forecasting.} A \texttt{statsmodels} ARIMA model forecasting unemployment-rate trajectories (24 prediction records, one-quarter horizon). This use case is not currently high-risk under Annex III but is representative of central-bank and macroeconomic-research pipelines where SR 11-7 / Bank of International Settlements model-risk obligations apply.
\end{itemize}

For each, we ran the instrumentation, applied the same five-use-case battery as the vol-forecasting demonstration, and verified that the chain-integrity, replay, drift, validation, export, and RTE capabilities all execute without code change. \Cref{tab:crossdomain} summarises the cross-domain results.

\begin{table}[H]
\centering
\caption{Cross-domain instrumentation: same package, same schema, same use-case battery, three distinct ML pipelines. The credit-classifier example exercises an EU AI Act Annex III high-risk use case directly; the macro-forecaster exercises an SR 11-7-relevant time-series use case.}\label{tab:crossdomain}
\footnotesize
\begin{tabular}{@{}p{3.0cm}rrlll@{}}
\toprule
Pipeline & Records & Models & Library & Use case 1-5 status & Annex III ref. \\
\midrule
Vol forecasting (Khan 2026) & 6{,}825 & 7 & arch, lightgbm, xgboost & 5/5 pass & not in scope \\
Credit scoring & 800 & 1 & scikit-learn & 5/5 pass & §5(b) high-risk \\
Macro (unemployment) & 24 & 1 & statsmodels & 5/5 pass & not in scope \\
\bottomrule
\end{tabular}
\end{table}

Two observations support the schema's generality. \emph{First, the integration cost is essentially identical across the three pipelines}: the \texttt{@auditable} decorator wraps the prediction function in each case, with feature-name and data-source metadata supplied through the same caller-supplied arguments. No domain-specific schema extension was required. \emph{Second, the same chain-integrity, replay, and export properties hold across pipelines.} The credit-classifier and macro logs both verify cleanly via \texttt{verify\_chain}, replay bit-identically for the inference function, and produce regulator-export bundles whose integrity checks pass identically to the vol-forecasting case.

The cross-domain validation is intentionally modest in scope: the credit-classifier uses synthetic data with \texttt{make\_classification}, and the macro forecaster uses a synthetic AR(1)-with-seasonality DGP. We do not claim to have validated the package against a production bank credit-scoring system. We claim only that the schema and the package transfer without modification across ML libraries and ML problem types within the limits of the demonstrated examples. The \texttt{examples/} directory carries both pipelines under open license for independent reproduction.

\subsection{Adversarial robustness: detection rates under tampering}\label{sec:demonstration_adversarial}

The integrity claims in \Cref{sec:demonstration_replay} and \Cref{sec:demonstration_export} require empirical support beyond ``the chain verifies on the unmodified log.'' This subsection reports controlled tampering experiments. For each of four attack scenarios, we copy the 6,825-record log, apply the tampering, and run \texttt{verify\_chain} from the package's public API. \Cref{tab:adversarial} reports detection.

\begin{table}[H]
\centering
\caption{Detection rates under controlled tampering. The baseline is the unmodified 6{,}825-record demonstration log. ``Detected'' means \texttt{verify\_chain} returns \texttt{valid=False} with at least one explanatory error after the tampering is applied. The full per-scenario detail (including the tampered records' identifiers and the verification error messages) is in \texttt{studies/results/adversarial\_summary.json}.}\label{tab:adversarial}
\footnotesize
\begin{tabular}{@{}p{5.2cm}p{6.0cm}cc@{}}
\toprule
Scenario & Tamper applied & Detected & Error count \\
\midrule
\textit{Baseline} (no tampering) & --- & valid & 0 \\
\midrule
A. Single-record value corruption & Change \texttt{prediction.value} of one record from 0.1406 to 99.99 inside the JSON column & \checkmark yes & 1 \\
B. Chain-link tampering & Replace embedded \texttt{prev\_record\_hash} of one record with 64-zero hash & \checkmark yes & 1 \\
C. Tail truncation & Delete the last 100 records (chain still self-consistent on remaining 6{,}725) & \textbf{no} & 0 \\
D. Middle-record deletion & Delete one record from the middle of the chain (id=3500) & \checkmark yes & 1 \\
\bottomrule
\end{tabular}
\end{table}

Three observations matter for the integrity-protection claim.

\textit{First, scenarios A, B, and D are all detected.} The hash chain catches any tampering that changes the content of a single record (A), breaks the link between two consecutive records (B), or removes a record from the interior of the chain (D). The detection mechanism in each case is the chain's structural invariant: \texttt{record\_id = SHA-256(canonical-JSON content)} for every record AND \texttt{prev\_record\_hash = record\_id of immediately preceding record}. Any mutation of either field is mechanically caught by \texttt{verify\_chain}.

\textit{Second, scenario C is the honest negative result.} A tail-truncation attack removes a contiguous block of trailing records but leaves the remaining sub-chain self-consistent. \texttt{verify\_chain} sees valid records 1 through 6,725 forming an unbroken hash chain ending at the truncated boundary, with no internal evidence that records 6,726--6,825 ever existed. Detecting this attack requires \emph{external anchoring}: periodic publication of the latest \texttt{record\_id} to a trusted third-party store (an internal audit ledger, a public blockchain, a regulator-controlled timestamp service, or signed publication of latest-record manifest by an HSM) so that an auditor can verify the present log's latest record matches the externally-anchored expected value. We do not provide external attestation in v0.1.0; we provide the substrate on which a firm can layer its preferred attestation strategy. \Cref{sec:future} lists external-attestation hooks as the v0.2.0 priority.

\textit{Third, the detection guarantee is conditional on the verifier reading the chain itself, not a derived summary.} If a regulator accepts a chain-verification summary (``the firm reports verify\_chain returned valid'') without running the verification themselves, the firm could in principle present a tampered chain with a fabricated ``valid'' attestation. The regulator-export bundle (\Cref{sec:demonstration_export}) addresses this by making the verification trivially executable from the bundle alone: a regulator with only \texttt{sha256sum} and a JSON parser can independently re-derive the chain's validity without trusting any attestation from the firm.

\subsection{Summary of the demonstration}\label{sec:demonstration_summary}

The five use cases together demonstrate that a 6,825-record audit trail produced by \texttt{mr-audit} on a real seven-model volatility-forecasting pipeline can be:

\begin{itemize}[leftmargin=*]
    \item Replayed exactly (use case 1, 5/5 sampled records match bit-identically),
    \item Monitored for distributional drift across calendar windows (use case 2, 44/44 months flagged consistently with the documented regime shift) on both inputs and outputs (\texttt{detect\_drift}, \texttt{detect\_output\_drift}),
    \item Used as the sole basis for a per-model validation report (use case 3, 7 models, 975 predictions each),
    \item Exported as a regulator-portable, integrity-tied bundle (use case 4, 841 KB, all checks pass),
    \item Used as the data source for an audit-grade right-to-explanation reconstruction of any historical prediction (use case 5).
\end{itemize}

The demonstration's purpose is not to argue that the seven-model panel itself satisfies SR 11-7 (that determination is the bank's, not the package's); it is to show that the technical record-keeping that an SR 11-7-compliant deployment would require is producible and verifiable using vendor-neutral, open-source tooling at a realistic data scale.
